---
title: "EDA"
format:
  html:
    toc: true
    html-math-method: katex
    css: styles.css
editor: visual
---

## 

## Import libraries

```{r}
library(tidyverse)
library(GGally)
library(gender)
```

```{r}
# automated EDA libraries
library(SmartEDA)
library(DataExplorer)
library(dlookr)
library(skimr)
```

 

## Read in data

```{r}
CustomerDemographic <- read.csv('cleansed data/CustomerDemographic.csv')
CustomerAddress <- read.csv('cleansed data/CustomerAddress.csv')
Transactions <- read.csv('cleansed data/Transactions.csv')
NewCustomerList <- read.csv('cleansed data/NewCustomerList.csv')

Customer_U_gender <- read.csv('cleansed data/subsets/Customer_U_gender.csv')
```

## Convert data types

```{r}
skim(CustomerDemographic)
str(CustomerDemographic)
```

We won't need first_name and last_name for training.

However, we need names to present our findings to business people.

```{r}
CustomerDemographic$customer_id <- as.character(CustomerDemographic$customer_id)
class(CustomerDemographic$customer_id)
```

```{r}
CustomerDemographic$gender <- factor(CustomerDemographic$gender,
                                        ordered = FALSE)
class(CustomerDemographic$gender)
```

```{r}
CustomerDemographic$job_title <- factor(
  CustomerDemographic$job_title,
  ordered = FALSE)

CustomerDemographic$job_industry_category <- factor(
  CustomerDemographic$job_industry_category,
  ordered = FALSE)
# levels(CustomerDemographic$job_industry_category)
# to list out levels in a factor

CustomerDemographic$wealth_segment <- factor(
  CustomerDemographic$wealth_segment,
  ordered = TRUE,
  levels = c('Mass Customer', 'Affluent Customer', 'High Net Worth'))


CustomerDemographic$owns_car <- factor(
  CustomerDemographic$owns_car,
  ordered = FALSE)
```

```{r}
str(CustomerDemographic$DOB)

CustomerDemographic$DOB <- as.Date(CustomerDemographic$DOB)


head(CustomerDemographic$DOB)
sum(is.na(CustomerDemographic$DOB)) # still expect 87 NAs
```

```{r}
str(CustomerDemographic)
```

First impression and to-do list

-   change customer_id to chr type

-   recode gender, job_title, job_industry_category, wealth_segment, owns_car to factor

-   recode DOB to date

-   last_name's missing values are not needed to impute, leave as is

-   impute job_titles and job_industry_category

-   87 missing for DOB, customers with "U" as gender, and tenure (likely the same 87 customers)

-   reduce cardinality of job_industry_category by generalizing (Doctor, Engineer, Coach, Designer, etc.)

```{r}
skim(CustomerAddress)
str(CustomerAddress)
```

```{r}
CustomerAddress$customer_id <- as.character(CustomerAddress$customer_id)
CustomerAddress$postcode <- as.character(CustomerAddress$postcode)
```

```{r}
skim(Transactions)
str(Transactions)
```

```{r}
Transactions$transaction_id <- as.character(Transactions$transaction_id)
Transactions$product_id <- as.character(Transactions$product_id)
Transactions$customer_id <- as.character(Transactions$customer_id)

Transactions$transaction_date <- as.Date(Transactions$transaction_date)
Transactions$online_order <- factor(Transactions$online_order, ordered = FALSE)
Transactions$order_status <- factor(Transactions$order_status, ordered = FALSE)
Transactions$product_line <- factor(Transactions$product_line, ordered = FALSE)

Transactions$product_first_sold_date <- as.Date(Transactions$product_first_sold_date)
```

```{r}
Transactions$product_class <- factor(Transactions$product_class, 
                                     ordered = TRUE,
                                     levels = c('low', 'medium', 'high'))
Transactions$product_size <- factor(Transactions$product_size, 
                                    ordered = TRUE,
                                    levels = c('small', 'medium', 'large'))
Transactions$brand <- factor(Transactions$brand, ordered = FALSE)
```

```{r}
str(Transactions)
```

```{r}
skim(NewCustomerList)
str(NewCustomerList)
```

```{r}
# convert gender, job_title, job_industry_category, wealth segment, owns_car, postcode, state, country to factor type
NewCustomerList$gender <- as.factor(NewCustomerList$gender)
NewCustomerList$job_title <- as.factor(NewCustomerList$job_title)

NewCustomerList$job_industry_category <- as.factor(NewCustomerList$job_industry_category)
NewCustomerList$wealth_segment <- factor(
  NewCustomerList$wealth_segment,
  ordered = TRUE,
  levels = c('Mass Customer', 'Affluent Customer', 'High Net Worth'))

NewCustomerList$owns_car <- as.factor(NewCustomerList$owns_car)
NewCustomerList$postcode <- as.character(NewCustomerList$postcode)
NewCustomerList$state <- as.factor(NewCustomerList$state)
NewCustomerList$country <- as.factor(NewCustomerList$country)
```

```{r}
str(NewCustomerList)
```

## Impute or remove data

### Transactions

```{r}
# subset transactions with missing product info
Tx_missing_prodInfo <- Transactions[is.na(Transactions$product_first_sold_date),]
```

```{r}
plot_na_pareto(Transactions) +
  ylim(0,500)
```

Can we tell whether missing values in these different columns belong to the same observations?

```{r}

```

### gender

Since first names often indicate gender, it should be possible to predict gender using names.

We will impute gender (or more precisely, sex assigned at birth) based on first names using historical data.

The R package [*gender*](https://www.rdocumentation.org/packages/gender/versions/0.6.0) uses historical datasets from the U.S. Social Security Administration, the U.S. Census Bureau (via [IPUMS USA](https://usa.ipums.org/usa/)), and the [North Atlantic Population Project](https://www.nappdata.org/napp/) to provide predictions of gender for first names for particular countries and time periods.

Other package/API options

-   <https://gender-api.com/>

-   [genderize.io](genderize.io)

```{r}
# one limitation of using gender package instead of genderize is
# the ability to specify country = 'AU'

# The "genderize" method uses the Genderize.io <https://genderize.io/> API, 
# which is based on "user profiles across major social networks."
imputedUCustomers <- gender(Customer_U_gender$first_name, 
                            method = "genderize")
imputedUCustomers
```

check for missing values where gender() couldn't impute.

```{r}
sum(is.na(imputedUCustomers$gender))

imputedUCustomers |> filter(is.na(gender) == TRUE)
```

Manual imputation of these three names (source: thebump.com)

[Goldarina female name](https://www.thebump.com/b/goldarina-baby-name)

[Devland male name](https://www.thebump.com/b/devland-baby-name)

[Sutherlan male name](https://www.thebump.com/b/sutherland-baby-name)

```{r}
imputedUCustomers <- imputedUCustomers |>
  mutate(gender = case_when(
    name == "Goldarina" ~ "F",
    name == "Devland" ~ "M",
    name == "Sutherlan" ~ "M",
    TRUE ~ as.character(gender)
  ))
```

standardize gender column values

```{r}
imputedUCustomers <- imputedUCustomers |>
  mutate(gender = ifelse(gender == "male", "M", "F"))

imputedUCustomers
```

```{r}
str(imputedUCustomers)
str(CustomerDemographic)
```

apply imputed gender values to original CustomerDemographic dataset.

```{r}
imputedU_CD <- CustomerDemographic |> 
  left_join(imputedUCustomers, by = c("first_name" = "name")) |>
  # uses mutate and ifelse to update the "gender" column 
  # based on the values in the "gender.x" and "gender.y" columns.
  mutate(gender = ifelse(gender.x == "U", gender.y, gender.x)) |>
  select(-starts_with("gender."))

# Now, 'imputedU_CD' contains the updated/imputed 'gender' column

```

imputedU_CD should have same number of rows as CustomerDemographic, ie. 3997.

Check if majority of imputed U customers are Male.

```{r}
table(imputedUCustomers$gender)
```

Turns out our hypothesis that there are more IT male customers is wrong.

```{r}
table(
  imputedU_CD[imputedU_CD$job_industry_category == "IT", "gender"]
  )
```

The business attracts IT customers roughly equally in terms of gender.

```{r}
plot_bar(imputedU_CD, 
         by = 'gender', 
         ncol = 2)

plot_bar(CustomerDemographic, 
         by = 'gender', 
         ncol = 2)
```

```{r using-genderize}
# using genderize package

# x <- Customer_U_gender$first_name
# 
# xPrepared = textPrepare(x)
# 
# givenNames <- findGivenNames(xPrepared, 
#                             country = 'AU',
#                             textPrepare = FALSE,
#                             progress = FALSE)
# 
# genderize(x, 
#           genderDB = givenNames,
#           progress = TRUE)
```

```{r}
CustomerDemographic |> 
  filter(gender == "U")

summary(CustomerDemographic |> 
  filter(gender == "U"))
```

U gender customers also don't have info on DOB and tenure

### online_order

```{r}

```

## Join data

```{r}
customerInfo <- merge(CustomerDemographic, CustomerAddress,
                      by = 'customer_id')
```

```{r}
customerInfo.and.Tx <- merge(customerInfo, Transactions,
                             by = 'customer_id')
```

## Label data (define target)

## Split data into train and test

## Explore data

Frequency distribution of factor variables by count.

4 columns ignored with more than 50 categories.

first_name: 3138 categories

last_name: 3723 categories

DOB: 3446 categories

job_title: 196 categories

```{r}
summary(CustomerDemographic$past_3_years_bike_related_purchases)
boxplot(CustomerDemographic$past_3_years_bike_related_purchases, horizontal = TRUE)
```

```{r}
# from DataExplorer

# plot factor variables based on frequency (number of customers in each category)
plot_bar(CustomerDemographic, 
         nrow = 2, ncol = 2)

# with indicates name of continuous feature to be summed. 
# default is NULL, i.e., frequency.
plot_bar(CustomerDemographic, 
         with = "past_3_years_bike_related_purchases",
         nrow = 2, ncol = 2)
```

Gender distribution is pretty balanced, so we will not have any gender bias in training.

Most of the customers in 2017 are in the Mass segment and work in Manufacturing, Financial Services, and Health.

This majority of the customers also have the most bike related purchases in the last 3 years.

```{r}
# Assuming your data frame is called CustomerDemographic
CustomerDemographic |>
  group_by(job_industry_category) |>
  summarise(total_purchases = sum(past_3_years_bike_related_purchases, na.rm = TRUE)) |>
  arrange(desc(total_purchases)) |>
  ggplot(aes(x = reorder(job_industry_category, total_purchases), 
             y = total_purchases, 
             by = job_industry_category)) +
  geom_bar(stat = "identity") +
  labs(title = "Bike Purchases by Job Industry Category",
       x = "Job Industry Category",
       y = "Total Past 3 Years Bike Related Purchases") +
  coord_flip()
```

```{r}
plot_bar(CustomerDemographic, 
         by = 'gender', 
         ncol = 2)
```

Customers who did not provide their gender information appear to work mostly in IT.

```{r}
nrow(
  CustomerDemographic |> 
    filter(job_industry_category == "IT",
           gender == "U")
)

nrow(
  CustomerDemographic |> 
    filter(gender == "U")
)
```

A [2020 study](https://anitab.org/research-and-impact/top-companies/2020-results/) by the AnitaB.org Institute found that women make up 28.8% of the tech workforce, a steady increase from the past few years \-- 25.9% in 2018 and 26.2% in 2019.

So we will assume/expect that majority of U gender customers are Male.

```{r}
# current gender distribution, before imputation
table(
  (CustomerDemographic |> 
     filter(job_industry_category == "IT"))[, "gender"])
```

```{r}
plot_bar(CustomerDemographic, 
         by = 'wealth_segment', 
         ncol = 2)
```

```{r}
plot_bar(CustomerDemographic, 
         by = 'job_industry_category', 
         ncol = 2)
```

## 

```{r}
# library(flextable)
dlookr::diagnose_numeric(CustomerDemographic)
```
